### 2022-08-30 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@d74d2d8fe84f81e9431ea1c135d8cb274a599339 🚀
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@e95f073e8afe0ab4deb3916811d8ca7d3cf74bd9 🚀

%
Norma\,Forero Rocio Munoz%
%
\,\textsuperscript{1,2,‡}\quad %
Renata L.\,Muylaert%
%
\,\textsuperscript{3}\quad %
Stephanie N.\,Seifert%
%
\,\textsuperscript{4}\quad %
Gregory F.\,Albery%
%
\,\textsuperscript{5}\quad %
Daniel J.\,Becker%
%
\,\textsuperscript{6}\quad %
Colin J.\,Carlson%
%
\,\textsuperscript{7,8,9,‡}\quad %
\href{https://orcid.org/0000-0002-0735-5184}{Timothée\,Poisot}%
%
\,\textsuperscript{1,2,‡}

@@

Disease emergence is complex, and is driven not only by animal-human
contact, but also by the underlying evolutionary dynamics in viral
reservoirs.\textsuperscript{1} Although host richness is often used as a
superficial proxy for spillover risk,\textsuperscript{2,3} these
approaches oversimplify the relevant interspecific heterogeneity in
immunology, behavior, and other traits, and therefore overlook unique
host pools that allow for the rapid evolution of highly divergent
viruses.\textsuperscript{4} In the case of generalist pathogens like
betacoronaviruses, there is conceptual and empirical support to the idea
that these community-level mechanisms are even more
important,\textsuperscript{5} particularly given that cross-species
transmission may, as a rule, structure viral evolution more than
co-divergence with hosts.\textsuperscript{6} This creates a disconnect
between coevolutionary theory and most existing ecological frameworks
for mapping spillover risk.

The geographic mosaic theory of coevolution (GMTC) attempts to
explicitly connect microevolutionary dynamics to the macroecology and
biogeography of symbiotic interactions.\textsuperscript{7} The GMTC
posits that coevolutionary processes among pairs\textsuperscript{8} or
complexes\textsuperscript{9} of species are structured in space by the
rippling effects of abiotic conditions onto evolutionary mechanisms,
giving rise to fragmented systems with different ecologies over large
spatial extents.\textsuperscript{10} The GMTC predicts a spatial
fragmentation of coevolutionary dynamics under the joint action of three
processes:\textsuperscript{11} coevolutionary hot- and coldspots, which
appear when the intensity of \emph{interaction} (in terms of reciprocal
fitness consequences) varies spatially; selection mosaics, wherein the
intensity of \emph{selection} varies across space, driven by both the
biotic complexity of the community (locally diverse hosts and viruses
are more biotically complex) and the local favorability of the
environment;\textsuperscript{12} and trait remixing, which occurs when
coevolutionary dynamics change when community-level \emph{functional
traits} change through meta-community dynamics.

@@

Overall, these results suggest that the boundaries of bat and
betacoronavirus biogeographic regions are largely consistent. This may
be surprising, given that cross-species transmission may play a stronger
role in coronavirus diversification than
cospeciation---\textsuperscript{2}a property that would theoretically
allow for substantial broad divergence in their biogeography. However,
host jumps at the family level or higher are relatively rare and
significant events in coronavirus evolutionary
history;\textsuperscript{2,13} as a result, the mosaic of
betacoronavirus phylogeography is assembled from a set of overlapping
smaller coevolutionary systems, superimposed in space and filtered by
the importance of different subgroups in local host communities. For
example, the most speciose and cosmopolitan family of bats, the vesper
bats (Vespertilionidae), are considered the primary hosts of the
subgenus \emph{Merbecovirus} (MERS-like viruses);\textsuperscript{3,13}
but in the Americas, where merbecoviruses are the only lineage present,
they have only been found in other bat taxa (\emph{e.g.}, Molossidae,
Phyllostomidae).\textsuperscript{14--17} At the coarsest scale, these
heterogeneities are lost, and betacoronavirus biogeography tracks the
deep rifts in bat evolutionary history---but within broad regions, the
component coevolutionary systems may have very different dynamics.

@@

Bats, the second most diverse groups of mammals, are found worldwide;
gradients in their species richness generally track broader patterns of
mammal diversity, with a striking Neotropical hotspot (especially in the
Amazon basin) and a secondary hotspot centered in Indochina. These
hotspots of bat diversity are generally presumed to be hotspots of viral
adaptive radiation, and therefore areas of concern for human
health.\textsuperscript{2,18} However, the hotspots of known bat
betacoronavirus hosts show a distinct pattern, with primary hotspots
(both in terms of area and higher values) of host richness situated in
southeast Asia, parts of southern Europe, and to a lesser extent parts
of Africa in the -25-0 range of latitudes (fig.~\ref{fig:richness};
top). Although hundreds of species likely host undiscovered
betacoronaviruses, machine learning predictions have suggested that
these undiscovered reservoirs should follow the same diversity
gradient.\textsuperscript{19} In principle, these hotspots of
locally-diverse, virus-rich bat communities should drive more adaptive
diversification in their viruses.

@@

However, we find that the global pattern of betacoronavirus phylogenetic
distinctiveness is quite distinct from both bat host richness and
phylogenetic distinctiveness (fig.~\ref{fig:richness}; bottom). In
contrast to the sparsity of Neotropical betacoronavirus hosts, South and
Central America have the most evolutionary distinct hosts \emph{and}
viruses, followed by secondary hotspots in southeast Asia and the Rift
Valley region have mostly distinct viruses. Some degree of sampling bias
may contribute to these patterns: for example, the Neotropics are one of
the places where the fewest bat betacoronavirus sequences have been
generated,\textsuperscript{20--22} resulting in a sparser phylogenetic
tree, and artificially inflating distinctiveness; conversely,
disproportionate research effort in eastern China\textsuperscript{23}
may have led to a more complete inventory of the local diversity of
coronaviruses, again inflating these metrics relative to underlying
patterns. Even accounting for these potential biases, though, there is
obvious heterogeneity in betacoronavirus evolutionary distinctiveness
that is distinct from overall bat diversity.

Overall, these patterns recapitulate the evolutionary history of both
the order Chiroptera and the genus \emph{Betacoronavirus}. Horseshoe
bats (Rhinolophidae) include the reservoirs of the SARS-like viruses
(subgenus \emph{Sarbecovirus}), the group of pandemic threats that have
been of the greatest interest to researchers\textsuperscript{13} (and so
have been sampled most intensively).\textsuperscript{23} The hotspots of
host richness and viral diversity in southeast Asia---both of which are
disproportionately high, considering the global landscape of bat species
richness---are almost entirely driven by viral adaptive radiation
through host switching within this clade\textsuperscript{3,19}. In
contrast, the Neotropical hotspot of viral distinctiveness is driven by
isolation by host vicariance. Out of the four main groups of
betacoronaviruses, only merbecoviruses have been found in animals in the
Americas--- an introduction that is generally presumed to be
ancient.\textsuperscript{3,24} While comparatively understudied, New
World merbecoviruses have been found in the ghost-faced bats
(Mormoopidae), Neotropical leaf-nosed bats (Phyllostomidae), and
free-tailed bats (Molossidae).\textsuperscript{14--17} The former two
groups and a clade of the latter are endemic to the Neotropics, while
the explosive adaptive radiations of the phyllostomids are responsible
for the hotspot of bat diversity in the Amazon.\textsuperscript{25}
Together, these clades of New World bats play host to a distinct regime
of betacoronavirus coevolution.

@@

The existence of well-defined cophylogenetic regions suggests that the
bat-betacoronavirus system is spatially fragmented enough to create
divergent coevolutionary trajectories; in turn, this coevolutionary
mosaic may alter the risk of zoonotic emergence. These ideas are,
respectively, supported by the existence of hotspots of viral uniqueness
and the diverse origins of human betacoronaviruses. Together, this
framework points to a predictable relationship between host community
structure and coevolutionary pressure: phylogeographic structure in bat
hosts (and their diverse immune strategies)\textsuperscript{26} creates
a landscape of selective pressure; the trajectory of viruses'
coevolutionary response is, in turn, constrained by their opportunities
for either specialization or diversification through host jumps and
recombination.

@@

Compared to approaches that map emergence risk based only on the number
of known bat hosts of betacoronaviruses, our framework suggests regions
where high viral sharing dominates coevolutionary dynamics---such as
Latin America, or Eurasia above a northing of 30---would pose less of a
relative risk of zoonotic emergence. Nevertheless, areas of high host
uniqueness coupled with high viral sharing (red-to-pink in
fig.~\ref{fig:trivariate}) could create hotspots facilitated by viral
codivergence. Our framework identifies Madagascar, where most bat
species are endemic following evolutionary divergence from sister
species in both African and Asian continents,\textsuperscript{27} as one
such hotspot; interestingly, a recent study\textsuperscript{28} reported
a novel and highly divergent lineage of nobecoviruses from
Madagascar-endemic pteropid bat species (\emph{Pteropus rufus} and
\emph{Rousettus madagascariensis}), again supporting the predictive
power of the coevolutionary framework.

@@

The relationship between the underlying pathogen pool and emergence risk
is mediated by both human-wildlife interfaces (the probability of
spillover) and opportunities for onward horizontal transmission (the
probability that spillovers become epidemics)\textsuperscript{1}. As a
proxy for both, we finally overlaid the risk component from the
composite map (see above) with the proportion of built land, as a proxy
for a mix of habitat disturbance, potential for bat synanthropy or
contact with bridge hosts like livestock,\textsuperscript{29,30} and
human population density and connectivity\textsuperscript{1,31,32}
(fig.~\ref{fig:compound}). Accounting for these factors, most of South
America and Europe are at comparatively lower risk, as--although densely
populated--settlements tend to be in areas with lower potential risk.
Conversely, regions like Malaysia and the northern coast of Australia
have a high evolutionary risk component, but should represent a
relatively lower effective risk due to low human density. However,
southeast Asia, the Indian subcontinent, and scattered hotspots in
sub-Saharan Africa are at high risk due to the overlap between human
populations and natural opportunities for cross-species transmission of
betacoronaviruses.

@@

Reassuringly, these predictions correspond to the geographic origins of
the three bat-origin coronaviruses that have recently emerged in human
populations. While available information puts the spillover of
SARS-CoV-2 in a live animal market in Wuhan, China, the ultimate origin
of the virus is almost certainly in a divergent lineage of
sarbecoviruses from Indochina that was poorly characterized prior to the
pandemic.\textsuperscript{20--22} Similarly, the SARS-CoV outbreak began
in Guangdong province in 2002, reaching humans through small carnivore
bridge hosts, but was eventually traced back to a set of likely
progenitor viruses found in cave-dwelling horseshoe bats in Yunnan
province;\textsuperscript{33} nearby, antibody evidence has indicated
human exposure to SARS-like viruses.\textsuperscript{34} MERS-CoV was
first detected in Jordan, but is widespread in camels in East Africa and
the Middle East, and may have reached its bridge host decades earlier
than originally supposed;\textsuperscript{35} as a result, the geography
of the original bat-to-camel transmission is still widely regarded as
uncertain. All of these are broadly consistent with the risk factors we
identify. Notably, India and west Africa are additional hotspots that
have yet to experience the emergence of a bat coronavirus into human
populations, but may still be at risk---particularly given known gaps in
bat surveillance,\textsuperscript{23} and a dense population in both
regions with global connectivity. In any of these regions, surveillance
on viral reservoirs can be paired with targeted monitoring of high-risk
human populations (i.e., those with regular wildlife
contact)\textsuperscript{36} for maximum impact.

@@

Bats emerged around 64 million years ago, and are one of the most
diverse mammalian orders, with more than 1,400 estimated
species.\textsuperscript{37,38} They exhibit a broad variety of habitat
use, behaviour, and feeding strategies, putting them at key positions in
the delivery and provisioning of several ecosystem services, tied to
important ecosystem-derived benefits to humans.\textsuperscript{39} Over
two-thirds of bats are know to be either obligate or facultative
insectivores, therefore actively contributing for agricultural pest
control,\textsuperscript{40,41} and vectors of pathogens that put a risk
on human health;\textsuperscript{42,43} some other species are essential
links in many seed-dispersal networks.\textsuperscript{44} However, many
of these species face a high risk of extinction, particularly given
persecution and killings that sometimes follows from messaging about
their role in disease emergence. Areas where bats, viruses, and humans
co-occur are not always hotspots of risk for human heath; as such,
developing more precise ways to map zoonotic hazards can help bats and
humans coexist safely, and support the conservation of these important
and unique animals.

Here, we propose a simple framework with broad explanatory power that
helps contextualize discoveries like highly divergent nobecoviruses in
Madagascar and the once-neglected adaptive radiation of sarbecoviruses
in the Indochinese peninsula. In doing so, it advances ecological theory
beyond the current state of the art for global maps of emergence risk.
For example, previous studies that have used host richness as a proxy
have predicted a high diversity of unsampled bat
viruses,\textsuperscript{18} bat coronaviruses,\textsuperscript{2} and
even specifically betacoronaviruses\textsuperscript{19} in both the
Amazon and southeast Asia. While we find that both regions are
characterized by unique and diverse communities of both hosts and
viruses, our framework is able to identify key differences between the
two systems. We find that the merbecovirus complex in Latin America has
been a unique branch of evolution separate from the rest of the global
pool, but with limited potential for viral diversification--- a finding
that is supported by previous work indicating a higher rate of
codivergence in Latin America.\textsuperscript{2,45} In contrast, in
southeast Asia, host richness and viral distinctiveness are high but
sharing is low; this suggests a different type of evolutionary dynamics
that could generate high local diversity of viruses through host
switching and viral recombination (see \emph{e.g.},\textsuperscript{13}
as well as the discovery of recombinant viruses with genetic material
from both the SARS-CoV and SARS-CoV-2 branches of the Sarbecovirus
lineage).\textsuperscript{46}

Both of these regions are priority areas for sampling, especially given
predictions that they contain many bat hosts of undiscovered
betacoronaviruses.\textsuperscript{19,23} However, both the evolutionary
and ecological aspects of emergence risk are higher in southeast
Asia---a fact that will only become more relevant, as bats track
shifting climates and exchange viruses with other species, creating a
hotspot of elevated cross-species transmission unique to the
region.\textsuperscript{31,47} Bats---and the spillover of their
viruses---are also sensitive to anthropogenic factors others than
climate change, including deforestation and other kinds of habitat loss,
increased stress, and greater contact with potential bridge hosts like
domesticated species.\textsuperscript{29,48--50} This represents a
challenge for both conservation strategies and pandemic
prevention,\textsuperscript{51} but identifying areas at risk, and
protecting the health of bats and ecosystems within those zones, can be
a win-win intervention for both.\textsuperscript{52--54}

@@

We downloaded the data on bats hosts of \emph{Betacoronavirus} from
\texttt{https://www.viralemergence.org/betacov} on
Apr.~2022,\textsuperscript{19} and filtered it to ``known'' hosts
(established before the emergence of SARS-CoV-2) and ``novel'' hosts
(confirmed through sampling and competence assays since the initial data
collection). The original database was assembled by a combination of
data mining and literature surveys, including automated alerts on the
``bats'' and ``coronavirus'' keywords to identify novel empirical
evidence of bats-betacoronaviruses associations; this yielded a total of
126 known hosts, 47 of which were novel hosts.

@@

We downloaded the rangemap of every current bat species that was
classified as an empirically documented host of \emph{Betacoronavirus}
from the previous step, according to recent IUCN
data.\textsuperscript{55} The range maps were subsequently rasterized
using the \texttt{rasterize} function from
\texttt{GDAL}\textsuperscript{56} at a resolution of approximately
100kmx100km at the equator. For every pixel in the resulting raster
where at least one bat host of \emph{Betacoronavirus} was present, we
extract the species pool (list of all known bat hosts), which was used
to calculate the following risk assessment components: bat phylogenetic
diversity, bat compositional uniqueness, and predicted viral sharing
risk.

@@

For all bat hosts of \emph{Betacoronavirus}, we extracted their
predicted viral sharing network, generated from a previously published
generalized additive mixed model of virus sharing by a tensor function
of phylogenetic distance and geographic range overlap across
mammals.\textsuperscript{62} This network stores pairwise values of
viral community similarity. To project viral sharing values into a
single value for every pixel, we averaged the pairwise scores. High
values of the average sharing propensity means that this specific extant
bat assemblage is likely to be proficient at exchanging viruses.

@@

To visualize the aggregated risk at the global scale, we combine the
three individual risk components (phylogenetic diversity, compositional
uniqueness, and viral sharing) using an additive color
model.\textsuperscript{63} In this approach, every risk component gets
assigned a component in the RGB color model (phylogenetic diversity is
green, compositional uniqueness is red, and viral sharing is blue). In
order to achieve a valid RGB measure, all components are re-scaled to
the {[}0,1{]} interval, so that a pixel with no sharing, no phylogenetic
diversity, and no compositional uniqueness is black, and a pixel with
maximal values for each is white. This additive model conveys both the
intensity of the overall risk, but also the nature of the risk as colors
diverge towards combinations of values for three risk components. Out of
the possible combinations, the most risky in terms or rapid
diversification and spillover potential is high phylogenetic diversity
and low viral sharing,\textsuperscript{64} in that this allows multiple
independent host-virus coevolutionary dynamics to take place in the same
location. In the colorimetric space, this correspond to yellow --
because the HSV space is more amenable to calculations for feature
extraction,\textsuperscript{65} we measured the risk level by
calculating the angular distance of the hue of each pixel to a reference
value of 60 (yellow), and weighted this risk level by the value
component. Specifically, given a pixel with colorimetric coordinates
\((h,s,v)\), its ranged weighted risk value is

@@

\leavevmode\hypertarget{ref-Agosta2010HowSpe}{}%
\CSLLeftMargin{4. }
\CSLRightInline{Agosta, S. J., Janz, N. \& Brooks, D. R. How specialists
can be generalists: Resolving the {`parasite paradox'} and implications
for emerging infectious disease. \emph{Zoologia (Curitiba)} \textbf{27},
151--162 (2010).}

\leavevmode\hypertarget{ref-Power2004PatSpi}{}%
\CSLLeftMargin{5. }
\CSLRightInline{Power, A. G. \& Mitchell, C. E. Pathogen Spillover in
Disease Epidemics. \emph{The American Naturalist} \textbf{164}, S79--S89
(2004).}

>>>>

%
Norma R\,Forero-Muñoz%
%
\,\textsuperscript{1,2,‡}\quad %
Renata L.\,Muylaert%
%
\,\textsuperscript{3}\quad %
Stephanie N.\,Seifert%
%
\,\textsuperscript{4}\quad %
Gregory F.\,Albery%
%
\,\textsuperscript{5}\quad %
Daniel J.\,Becker%
%
\,\textsuperscript{6}\quad %
Colin J.\,Carlson%
%
\,\textsuperscript{7,8,9,‡}\quad %
\href{https://orcid.org/0000-0002-0735-5184}{Timothée\,Poisot}%
%
\,\textsuperscript{1,2,‡}

@@

Disease emergence is complex, and is driven not only by animal-human
contact, but also by the underlying evolutionary dynamics in viral
reservoirs.\textsuperscript{1} Although host richness is often used as a
superficial proxy for spillover risk,\textsuperscript{2--4} these
approaches oversimplify the relevant interspecific heterogeneity in
immunology, behavior, and other traits, and therefore overlook unique
host pools that allow for the rapid evolution of highly divergent
viruses.\textsuperscript{5} In the case of generalist pathogens like
betacoronaviruses, there is conceptual and empirical support to the idea
that these community-level mechanisms are even more
important,\textsuperscript{6} particularly given that cross-species
transmission may, as a rule, structure viral evolution more than
co-divergence with hosts.\textsuperscript{7} This creates a disconnect
between coevolutionary theory and most existing ecological frameworks
for mapping spillover risk.

The geographic mosaic theory of coevolution (GMTC) attempts to
explicitly connect microevolutionary dynamics to the macroecology and
biogeography of symbiotic interactions.\textsuperscript{8} The GMTC
posits that coevolutionary processes among pairs\textsuperscript{9} or
complexes\textsuperscript{10} of species are structured in space by the
rippling effects of abiotic conditions onto evolutionary mechanisms,
giving rise to fragmented systems with different ecologies over large
spatial extents.\textsuperscript{11} The GMTC predicts a spatial
fragmentation of coevolutionary dynamics under the joint action of three
processes:\textsuperscript{12} coevolutionary hot- and coldspots, which
appear when the intensity of \emph{interaction} (in terms of reciprocal
fitness consequences) varies spatially; selection mosaics, wherein the
intensity of \emph{selection} varies across space, driven by both the
biotic complexity of the community (locally diverse hosts and viruses
are more biotically complex) and the local favorability of the
environment;\textsuperscript{13} and trait remixing, which occurs when
coevolutionary dynamics change when community-level \emph{functional
traits} change through meta-community dynamics.

@@

Overall, these results suggest that the boundaries of bat and
betacoronavirus biogeographic regions are largely consistent. This may
be surprising, given that cross-species transmission may play a stronger
role in coronavirus diversification than
cospeciation---\textsuperscript{2}a property that would theoretically
allow for substantial broad divergence in their biogeography. However,
host jumps at the family level or higher are relatively rare and
significant events in coronavirus evolutionary
history;\textsuperscript{2,14} as a result, the mosaic of
betacoronavirus phylogeography is assembled from a set of overlapping
smaller coevolutionary systems, superimposed in space and filtered by
the importance of different subgroups in local host communities. For
example, the most speciose and cosmopolitan family of bats, the vesper
bats (Vespertilionidae), are considered the primary hosts of the
subgenus \emph{Merbecovirus} (MERS-like viruses);\textsuperscript{3,14}
but in the Americas, where merbecoviruses are the only lineage present,
they have only been found in other bat taxa (\emph{e.g.}, Molossidae,
Phyllostomidae).\textsuperscript{15--18} At the coarsest scale, these
heterogeneities are lost, and betacoronavirus biogeography tracks the
deep rifts in bat evolutionary history---but within broad regions, the
component coevolutionary systems may have very different dynamics.

@@

Bats, the second most diverse groups of mammals, are found worldwide;
gradients in their species richness generally track broader patterns of
mammal diversity,\textsuperscript{\textbf{Tanalgo2022Mapping?}} with a
striking Neotropical hotspot (especially in the Amazon basin) and a
secondary hotspot centered in Indochina. These hotspots of bat diversity
are generally presumed to be hotspots of viral adaptive radiation, and
therefore areas of concern for human health.\textsuperscript{2,19}
However, the hotspots of known bat betacoronavirus hosts show a distinct
pattern, with primary hotspots (both in terms of area and higher values)
of host richness situated in southeast Asia, parts of southern Europe,
and to a lesser extent parts of Africa in the -25-0 range of latitudes
(fig.~\ref{fig:richness}; top). Although hundreds of species likely host
undiscovered betacoronaviruses, machine learning predictions have
suggested that these undiscovered reservoirs should follow the same
diversity gradient.\textsuperscript{20} In principle, these hotspots of
locally-diverse, virus-rich bat communities should drive more adaptive
diversification in their viruses.

@@

However, we find that the global pattern of betacoronavirus phylogenetic
distinctiveness is quite distinct from both bat host richness and
phylogenetic distinctiveness (fig.~\ref{fig:richness}; bottom). In
contrast to the sparsity of Neotropical betacoronavirus hosts, South and
Central America have the most evolutionary distinct hosts \emph{and}
viruses, followed by secondary hotspots in southeast Asia and the Rift
Valley region have mostly distinct viruses. Some degree of sampling bias
may contribute to these patterns: for example, the Neotropics are one of
the places where the fewest bat betacoronavirus sequences have been
generated,\textsuperscript{21--23} resulting in a sparser phylogenetic
tree, and artificially inflating distinctiveness; conversely,
disproportionate research effort in eastern China\textsuperscript{24}
may have led to a more complete inventory of the local diversity of
coronaviruses, again inflating these metrics relative to underlying
patterns. Even accounting for these potential biases, though, there is
obvious heterogeneity in betacoronavirus evolutionary distinctiveness
that is distinct from overall bat diversity.

Overall, these patterns recapitulate the evolutionary history of both
the order Chiroptera and the genus \emph{Betacoronavirus}. Horseshoe
bats (Rhinolophidae) include the reservoirs of the SARS-like viruses
(subgenus \emph{Sarbecovirus}), the group of pandemic threats that have
been of the greatest interest to researchers\textsuperscript{14} (and so
have been sampled most intensively).\textsuperscript{24} The hotspots of
host richness and viral diversity in southeast Asia---both of which are
disproportionately high, considering the global landscape of bat species
richness---are almost entirely driven by viral adaptive radiation
through host switching within this clade\textsuperscript{3,20}. In
contrast, the Neotropical hotspot of viral distinctiveness is driven by
isolation by host vicariance. Out of the four main groups of
betacoronaviruses, only merbecoviruses have been found in animals in the
Americas--- an introduction that is generally presumed to be
ancient.\textsuperscript{3,25} While comparatively understudied, New
World merbecoviruses have been found in the ghost-faced bats
(Mormoopidae), Neotropical leaf-nosed bats (Phyllostomidae), and
free-tailed bats (Molossidae).\textsuperscript{15--18} The former two
groups and a clade of the latter are endemic to the Neotropics, while
the explosive adaptive radiations of the phyllostomids are responsible
for the hotspot of bat diversity in the Amazon.\textsuperscript{26}
Together, these clades of New World bats play host to a distinct regime
of betacoronavirus coevolution.

@@

The existence of well-defined cophylogenetic regions suggests that the
bat-betacoronavirus system is spatially fragmented enough to create
divergent coevolutionary trajectories; in turn, this coevolutionary
mosaic may alter the risk of zoonotic emergence. These ideas are,
respectively, supported by the existence of hotspots of viral uniqueness
and the diverse origins of human betacoronaviruses. Together, this
framework points to a predictable relationship between host community
structure and coevolutionary pressure: phylogeographic structure in bat
hosts (and their diverse immune strategies)\textsuperscript{27} creates
a landscape of selective pressure; the trajectory of viruses'
coevolutionary response is, in turn, constrained by their opportunities
for either specialization or diversification through host jumps and
recombination.

@@

Compared to approaches that map emergence risk based only on the number
of known bat hosts of betacoronaviruses, our framework suggests regions
where high viral sharing dominates coevolutionary dynamics---such as
Latin America, or Eurasia above a northing of 30---would pose less of a
relative risk of zoonotic emergence. Nevertheless, areas of high host
uniqueness coupled with high viral sharing (red-to-pink in
fig.~\ref{fig:trivariate}) could create hotspots facilitated by viral
codivergence. Our framework identifies Madagascar, where most bat
species are endemic following evolutionary divergence from sister
species in both African and Asian continents,\textsuperscript{28} as one
such hotspot; interestingly, a recent study\textsuperscript{29} reported
a novel and highly divergent lineage of nobecoviruses from
Madagascar-endemic pteropid bat species (\emph{Pteropus rufus} and
\emph{Rousettus madagascariensis}), again supporting the predictive
power of the coevolutionary framework.

@@

The relationship between the underlying pathogen pool and emergence risk
is mediated by both human-wildlife interfaces (the probability of
spillover) and opportunities for onward horizontal transmission (the
probability that spillovers become epidemics)\textsuperscript{1}. As a
proxy for both, we finally overlaid the risk component from the
composite map (see above) with the proportion of built land, as a proxy
for a mix of habitat disturbance, potential for bat synanthropy or
contact with bridge hosts like livestock,\textsuperscript{30,31} and
human population density and connectivity\textsuperscript{1,32,33}
(fig.~\ref{fig:compound}). Accounting for these factors, most of South
America and Europe are at comparatively lower risk, as--although densely
populated--settlements tend to be in areas with lower potential risk.
Conversely, regions like Malaysia and the northern coast of Australia
have a high evolutionary risk component, but should represent a
relatively lower effective risk due to low human density. However,
southeast Asia, the Indian subcontinent, and scattered hotspots in
sub-Saharan Africa are at high risk due to the overlap between human
populations and natural opportunities for cross-species transmission of
betacoronaviruses.

@@

Reassuringly, these predictions correspond to the geographic origins of
the three bat-origin coronaviruses that have recently emerged in human
populations. While available information puts the spillover of
SARS-CoV-2 in a live animal market in Wuhan, China, the ultimate origin
of the virus is almost certainly in a divergent lineage of
sarbecoviruses from Indochina that was poorly characterized prior to the
pandemic.\textsuperscript{21--23} Similarly, the SARS-CoV outbreak began
in Guangdong province in 2002, reaching humans through small carnivore
bridge hosts, but was eventually traced back to a set of likely
progenitor viruses found in cave-dwelling horseshoe bats in Yunnan
province;\textsuperscript{34} nearby, antibody evidence has indicated
human exposure to SARS-like viruses.\textsuperscript{35} MERS-CoV was
first detected in Jordan, but is widespread in camels in East Africa and
the Middle East, and may have reached its bridge host decades earlier
than originally supposed;\textsuperscript{36} as a result, the geography
of the original bat-to-camel transmission is still widely regarded as
uncertain. All of these are broadly consistent with the risk factors we
identify. Notably, India and west Africa are additional hotspots that
have yet to experience the emergence of a bat coronavirus into human
populations, but may still be at risk---particularly given known gaps in
bat surveillance,\textsuperscript{24} and a dense population in both
regions with global connectivity. In any of these regions, surveillance
on viral reservoirs can be paired with targeted monitoring of high-risk
human populations (i.e., those with regular wildlife
contact)\textsuperscript{37} for maximum impact.

@@

Bats emerged around 64 million years ago, and are one of the most
diverse mammalian orders, with more than 1,400 estimated
species.\textsuperscript{38,39} They exhibit a broad variety of habitat
use, behaviour, and feeding strategies, putting them at key positions in
the delivery and provisioning of several ecosystem services, tied to
important ecosystem-derived benefits to humans.\textsuperscript{40} Over
two-thirds of bats are know to be either obligate or facultative
insectivores, therefore actively contributing for agricultural pest
control,\textsuperscript{41,42} and vectors of pathogens that put a risk
on human health;\textsuperscript{43,44} some other species are essential
links in many seed-dispersal networks.\textsuperscript{45} However, many
of these species face a high risk of extinction, particularly given
persecution and killings that sometimes follows from messaging about
their role in disease emergence. Areas where bats, viruses, and humans
co-occur are not always hotspots of risk for human heath; as such,
developing more precise ways to map zoonotic hazards can help bats and
humans coexist safely, and support the conservation of these important
and unique animals.

Here, we propose a simple framework with broad explanatory power that
helps contextualize discoveries like highly divergent nobecoviruses in
Madagascar and the once-neglected adaptive radiation of sarbecoviruses
in the Indochinese peninsula. In doing so, it advances ecological theory
beyond the current state of the art for global maps of emergence risk.
For example, previous studies that have used host richness as a proxy
have predicted a high diversity of unsampled bat
viruses,\textsuperscript{19} bat coronaviruses,\textsuperscript{2} and
even specifically betacoronaviruses\textsuperscript{20} in both the
Amazon and southeast Asia. While we find that both regions are
characterized by unique and diverse communities of both hosts and
viruses, our framework is able to identify key differences between the
two systems. We find that the merbecovirus complex in Latin America has
been a unique branch of evolution separate from the rest of the global
pool, but with limited potential for viral diversification--- a finding
that is supported by previous work indicating a higher rate of
codivergence in Latin America.\textsuperscript{2,46} In contrast, in
southeast Asia, host richness and viral distinctiveness are high but
sharing is low; this suggests a different type of evolutionary dynamics
that could generate high local diversity of viruses through host
switching and viral recombination (see \emph{e.g.},\textsuperscript{14}
as well as the discovery of recombinant viruses with genetic material
from both the SARS-CoV and SARS-CoV-2 branches of the Sarbecovirus
lineage).\textsuperscript{47}

Both of these regions are priority areas for sampling, especially given
predictions that they contain many bat hosts of undiscovered
betacoronaviruses.\textsuperscript{20,24} However, both the evolutionary
and ecological aspects of emergence risk are higher in southeast
Asia---a fact that will only become more relevant, as bats track
shifting climates and exchange viruses with other species, creating a
hotspot of elevated cross-species transmission unique to the
region.\textsuperscript{32,48} Bats---and the spillover of their
viruses---are also sensitive to anthropogenic factors others than
climate change, including deforestation and other kinds of habitat loss,
increased stress, and greater contact with potential bridge hosts like
domesticated species.\textsuperscript{30,49--51} This represents a
challenge for both conservation strategies and pandemic
prevention,\textsuperscript{52} but identifying areas at risk, and
protecting the health of bats and ecosystems within those zones, can be
a win-win intervention for both.\textsuperscript{53--55}

@@

We downloaded the data on bats hosts of \emph{Betacoronavirus} from
\texttt{https://www.viralemergence.org/betacov} on
Apr.~2022,\textsuperscript{20} and filtered it to ``known'' hosts
(established before the emergence of SARS-CoV-2) and ``novel'' hosts
(confirmed through sampling and competence assays since the initial data
collection). The original database was assembled by a combination of
data mining and literature surveys, including automated alerts on the
``bats'' and ``coronavirus'' keywords to identify novel empirical
evidence of bats-betacoronaviruses associations; this yielded a total of
126 known hosts, 47 of which were novel hosts.

@@

We downloaded the rangemap of every current bat species that was
classified as an empirically documented host of \emph{Betacoronavirus}
from the previous step, according to recent IUCN
data.\textsuperscript{56} The range maps were subsequently rasterized
using the \texttt{rasterize} function from
\texttt{GDAL}\textsuperscript{57} at a resolution of approximately
100kmx100km at the equator. For every pixel in the resulting raster
where at least one bat host of \emph{Betacoronavirus} was present, we
extract the species pool (list of all known bat hosts), which was used
to calculate the following risk assessment components: bat phylogenetic
diversity, bat compositional uniqueness, and predicted viral sharing
risk.

@@

For all bat hosts of \emph{Betacoronavirus}, we extracted their
predicted viral sharing network, generated from a previously published
generalized additive mixed model of virus sharing by a tensor function
of phylogenetic distance and geographic range overlap across
mammals.\textsuperscript{63} This network stores pairwise values of
viral community similarity. To project viral sharing values into a
single value for every pixel, we averaged the pairwise scores. High
values of the average sharing propensity means that this specific extant
bat assemblage is likely to be proficient at exchanging viruses.

@@

To visualize the aggregated risk at the global scale, we combine the
three individual risk components (phylogenetic diversity, compositional
uniqueness, and viral sharing) using an additive color
model.\textsuperscript{64} In this approach, every risk component gets
assigned a component in the RGB color model (phylogenetic diversity is
green, compositional uniqueness is red, and viral sharing is blue). In
order to achieve a valid RGB measure, all components are re-scaled to
the {[}0,1{]} interval, so that a pixel with no sharing, no phylogenetic
diversity, and no compositional uniqueness is black, and a pixel with
maximal values for each is white. This additive model conveys both the
intensity of the overall risk, but also the nature of the risk as colors
diverge towards combinations of values for three risk components. Out of
the possible combinations, the most risky in terms or rapid
diversification and spillover potential is high phylogenetic diversity
and low viral sharing,\textsuperscript{65} in that this allows multiple
independent host-virus coevolutionary dynamics to take place in the same
location. In the colorimetric space, this correspond to yellow --
because the HSV space is more amenable to calculations for feature
extraction,\textsuperscript{66} we measured the risk level by
calculating the angular distance of the hue of each pixel to a reference
value of 60 (yellow), and weighted this risk level by the value
component. Specifically, given a pixel with colorimetric coordinates
\((h,s,v)\), its ranged weighted risk value is

@@

\leavevmode\hypertarget{ref-Sanchez2022Strategy}{}%
\CSLLeftMargin{4. }
\CSLRightInline{Sánchez, C. A. \emph{et al.} A strategy to assess
spillover risk of bat SARS-related coronaviruses in southeast asia.
\emph{Nature communications} \textbf{13}, 1--12 (2022).}

\leavevmode\hypertarget{ref-Agosta2010HowSpe}{}%
\CSLLeftMargin{5. }
\CSLRightInline{Agosta, S. J., Janz, N. \& Brooks, D. R. How specialists
can be generalists: Resolving the {`parasite paradox'} and implications
for emerging infectious disease. \emph{Zoologia (Curitiba)} \textbf{27},
151--162 (2010).}
